R Notebook

### 2021-10-29 FEM_R/Wealth_Health_Socioeconomics.Rmd
Added code for generating replacement repl populations, along with some of the accounting needed and some imputation to get the population to run through the model (more needed before its working correctly)

require(tidyverse)
require(haven)
require(ggplot2)

# for multiple imputation replacement
require(mice)
require(data.table)

@@

### ACCOUNTING
# if smoken == 1 then most likely l2smoken == 1 also
if (is.na(test2$l2smoken)) {
  test2$l2smoken <- test2$smoken
}

# if l2smokev == 1 then smokev must also == 1
if (is.na(test2$smokev)) {
  test2$smokev <- test2$l2smokev
  # also, if l2smoken == 1 then l2smokev must also equal 1
  if (test2$l2smoken == 1) {
    test2$l2smokev <- 1
  }
  # same for smoken
  if (test2$smoken == 1)
    test2$l2smoken <- 1
}

# no variability in problem_drinker -> l2problem_drinker in median wealth group,
# so assume same here
if (is.na(test2$l2problem_drinker)) {
  test2$l2problem_drinker <- test2$problem_drinker
}

## OK!
## The test2 dataset now contains fully imputed data for quintile 1 based on
## values from quintile 3 so far. Can easily turn this into a function and 
## use different quintiles (or just do them without the function)
```

```{r}
## Step 2 is now to replace the values from the original repl dataset and save this as an alternative for running
# through the model

@@

```{r}
## FINALLY SAVE THE DATA
write_dta(repl.final, 'input_data/ELSA_repl_modified_quint1.dta')
```

>>>>

# Data manipulation and visualisation
require(tidyverse)
require(ggplot2)

# Handling Stata data files
require(haven)
require(foreign)

# for multiple imputation replacement
require(mice)

@@

### ACCOUNTING
# if smoken == 1 then most likely l2smoken == 1 also
# if (is.na(test2$l2smoken)) {
#   test2$l2smoken <- test2$smoken
# }
test2$l2smoken[is.na(test2$l2smoken)] <- test2$smoken

# if l2smokev == 1 then smokev must also == 1
# if (is.na(test2$smokev)) {
#   test2$smokev <- test2$l2smokev
#   # also, if l2smoken == 1 then l2smokev must also equal 1
#   if (test2$l2smoken == 1) {
#     test2$l2smokev <- 1
#   }
#   # same for smoken
#   if (test2$smoken == 1)
#     test2$l2smoken <- 1
# }
test2$smokev[is.na(test2$smokev)] <- test2$l2smokev
test2$l2smokev[test2$l2smoken == 1] <- 1
test2$l2smoken[test2$smoken == 1] <- 1

# no variability in problem_drinker -> l2problem_drinker in median wealth group,
# so assume same here
# if (is.na(test2$l2problem_drinker)) {
#   test2$l2problem_drinker <- test2$problem_drinker
# }
test2$l2problem_drinker[is.na(test2$l2problem_drinker)] <- test2$problem_drinker

## OK!
## The test2 dataset now contains fully imputed data for quintile 1 based on
## values from quintile 3 so far. Can easily turn this into a function and 
## use different quintiles (or just do them without the function)

rm(imp, test, test.1, test.3)
```

```{r}
## Step 2 is now to replace the values from the original repl dataset and save 
# this as an alternative for running through the model

@@

rm(quint1.replaced, repl.replaced, repl2)
```

Need to deal with some missing data and problems that are created by loading the data in R and saving to dta format.

Specifically, loading data in R loses the special missing characters that stata keeps and can deal with, just replaces with a simple (.) missing. One of the problem vars is educl (spouse education level), which is needed for the education imputation model. We can replace all missings with 0 as educl is used in the equation educl * married, which will be 0 if not married anyway.

## Whatever we do in here have to do to the baseline repl as well just to make 
# sure things are standardised

repl.final2 <- repl.final
repl2 <- repl

## handle missing educl data as described above
repl.final2$educl[is.na(repl.final2$educl)] <- 0
repl2$educl[is.na(repl$educl)] <- 0

table(repl.final2$educl)
summary(repl.final2$educl)
print('---------')

repl.final2$radadeduage[is.na(repl.final2$radadeduage)] <- 0
repl2$radadeduage[is.na(repl$radadeduage)] <- 0

repl.final2$ramomeduage[is.na(repl.final2$ramomeduage)] <- 0
repl2$ramomeduage[is.na(repl$ramomeduage)] <- 0

## USE THIS SPACE TO HOTDECK LOTS OF MISSING VARIABLES (start with lnly)
```

```{r}
## FINALLY SAVE THE DATA
write_dta(repl.final2, 'input_data/ELSA_repl_modified_quint1.dta', version = 12)
write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)
```
